---
title: "CHQA Analyst take-home task"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
```

Here is our first task:

---

## Part 1: Assemble the project cohort
The project goal is to identify patients seen for drug overdose, determine if they had an active opioid at the start of the encounter, and if they had any readmissions for drug overdose.

Your task is to assemble the study cohort by identifying encounters that meet the following criteria:

1. The patient’s visit is an encounter for drug overdose
2. The hospital encounter occurs after July 15, 1999
3. The patient’s age at time of encounter is between 18 and 35 (Patient is considered to be 35 until turning 36)

---

Sounds great. Let's start by taking a look at the data.

```{r}
allergies <- read_csv("datasets/allergies.csv")
allergies

encounters <- read_csv("datasets/encounters.csv")
encounters

medications <- read_csv("datasets/medications.csv")
medications

patients <- read_csv("datasets/patients.csv")
patients

procedures <- read_csv("datasets/procedures.csv")
procedures
```

Ok, we are chiefly interested in the `encounters` table, and basically want to filter it based on the specifications given in the task. Let's start by filtering the encounters by drug overdose. Looking at the data dictionary sheet for the `encounters` table, we can see that the `REASONCODE` column are SNOMED-CT codes.

We can lookup the code for a drug overdose here: https://browser.ihtsdotools.org/, which has the code as `55680006`.

```{r}
drug_overdoses <- filter(encounters, REASONCODE == 55680006)
drug_overdoses
```

Great, now we just need to filter for encounters that occur after July 15, 1999.

The `encounters` table has two column that represent the date of the encounter. `START` and `STOP`, further clarification would be neccessary to determine if the task is to find encounters that _begin_ after 07/15/1999 or _end_ at that date. For the purposes of this exercise, we'll go with encounters that _begin_ after that date due to the term _occur_ in the specification.

```{r}
after_date <- filter(drug_overdoses, START > "1999-07-15")
arrange(after_date, START)
```

Now we're concerned with encounters with patients between the ages of 18 and 35; we'll need to join the `patients` table to handle that.

```{r}
with_patients <- inner_join(after_date, patients, c("PATIENT" = "Id"))
with_patients
```

Based upon the wording in the specifications, the patient's age must be greater than or equal to 18 at the start of an encounter and less than or equal to 35 at the end of the encounter.

Let's make sure that there are no encounters in our table that has not ended, because a patient could age to 36 by the time the encounter is over.

```{r}
not_ended <- drop_na(with_patients, STOP)
not_ended
```

Turns out we're ok. Let's do the filtering now.
First we'll need to calculate the age of the patient at the start and end of the encounter.

```{r}
age <- mutate(not_ended, AGEATSTART = as.period(interval(BIRTHDATE, START))$year)
age <- mutate(age, AGEATSTOP = as.period(interval(BIRTHDATE, STOP))$year)
select(age, Id, AGEATSTART, AGEATSTOP)
```

```{r}
aged <- filter(age, AGEATSTART >= 18 & AGEATSTOP <= 35)
aged
```

That finishes up the first task.

---

### Part 2: Create additional fields
With your drug overdose encounter, create the following indicators:

1. `DEATH_AT_VISIT_IND`: `1` if patient died during the drug overdose encounter, `0` if the patient died at a different time
2. `COUNT_CURRENT_MEDS`: Count of active medications at the start of the drug overdose encounter
3. `CURRENT_OPIOID_IND`: `1` if the patient had at least one active medication at the start of the overdose encounter that is on the Opioids List (provided below), 0 if not 
4. `READMISSION_90_DAY_IND`: `1` if the visit resulted in a subsequent drug overdose readmission within 90 days, 0 if not 
5. `READMISSION_30_DAY_IND`: `1` if the visit resulted in a subsequent drug overdose readmission within 30 days, 0 if not overdose encounter, `0` if not
6. `FIRST_READMISSION_DATE`: The date of the index visit's first readmission for drug overdose. Field should be left as `N/A` if no readmission for drug overdose within 90 days

---

Opioids List:
* Hydromorphone 325Mg
* Fentanyl – 100 MCG
* Oxycodone-acetaminophen 100 Ml

---

Ok, looking at the data, it seems the only field we have to infer death on is in the `patients` table with the `DEATHDATE` column. If the date falls within the encounter dates, then we'll mark it `1`. The specifications don't state what to do if the patient hasn't died, this would need clarification, but for the purposes of this exercise we'll leave it blank in those cases.

```{r}
died <- mutate(aged, DEATH_AT_VISIT_IND = as.integer(DEATHDATE >= START & DEATHDATE <= STOP))
select(died, START, STOP, DEATHDATE, DEATH_AT_VISIT_IND)
```

For `COUNT_CURRENT_MEDS` we'll have to used the `medications` table.

For `CURRENT_OPOID_IND` we'll have to lookup the codes for the opoids in question; however, the codes in the `medications` table do not appear to match up with results found on: https://mor.nlm.nih.gov/RxNav/ (which is the RxNorm database the data dictionary mentioned). We would need clarification on this, but for this exercise we'll search by the `DESCRIPTION` column instead.

Some drugs have multiple components/ingredients. It's unsure whether we only should concern ourselves with the _pure_ drugs of interest or also these. For example: `Amlodipine 5 MG / Fentanyl 100 MCG / Olmesartan medoxomil 20 MG` vs `Fentanyl 100 MCG`. We would need clarification, but for the exercise we'll only examine the _pure_ drugs because multiple ingredients can modulate the effects of the drug in question, this is of course an assumption; and it's a little closer to the specification.

```{r}
opoids = c("Hydromorphone 325 MG", "Fentanyl 100 MCG", "Oxycodone-acetaminophen 100ML")

get_meds <- function(start, pt) {
  filter(medications, PATIENT == pt & START <= start & STOP <= start)
}

current_meds <- died %>%
  mutate(CURRENTMEDS = pmap(list(START, PATIENT), get_meds)) %>%
  mutate(COUNT_CURRENT_MEDS = map_int(CURRENTMEDS, nrow)) %>%
  mutate(CURRENT_OPOID_IND = map_int(CURRENTMEDS, function(med) any(med$DESCRIPTION %in% opoids)))
         
select(current_meds, CURRENTMEDS, COUNT_CURRENT_MEDS, CURRENT_OPOID_IND)
```

Now we can begin looking at readmissions. Let's calculate how many days it will be to the next encounter.

```{r}
readmission <- current_meds %>%
  arrange(START) %>%
  group_by(PATIENT) %>%
  mutate(READMISSION = as.period(lead(START) - STOP)$day) %>%
  mutate(FIRST_READMISSION_DATE = lead(START)) %>%
  ungroup

readmission$READMISSION
```

Great, now we can add the `READMISSION_90_DAY_IND` and `READMISSION_30_DAY_IND` columns.

```{r}
filtered_readmission <- readmission %>%
  mutate(READMISSION_90_DAY_IND = as.integer(READMISSION <= 90)) %>%
  mutate(READMISSION_30_DAY_IND = as.integer(READMISSION <= 30))

arrange(select(filtered_readmission, READMISSION, READMISSION_90_DAY_IND, READMISSION_30_DAY_IND), READMISSION)
```

Earlier when figuring out how many days there was till a readmission, we also put in what the readmission date is. We can use that here. This works because before doing that, we sorted the group dataframes by `START`. This way the next date is guaranteed to be the first readmission after a specific encounter.

```{r}
#first_readmission_dated <- mutate(filtered_readmission, FIRST_READMISSION_DATE = as_datetime(ifelse(READMISSION_90_DAY_IND == 1, READMISSIONDATE, NA)))
first_readmission_dated <- mutate(filtered_readmission, FIRST_READMISSION_DATE = replace(FIRST_READMISSION_DATE, READMISSION_90_DAY_IND == 0, NA))

arrange(select(first_readmission_dated, READMISSION, READMISSION_90_DAY_IND, READMISSION_30_DAY_IND, FIRST_READMISSION_DATE), READMISSION)
```

And now we can move onto the final task.

---

### Part 3: Export the data to a `CSV` file

Export a dataset containing these required fields:

| Field name                | Field Description                                                                                                                  | Data Type        |
| ------------------------- | ---------------------------------------------------------------------------------------------------------------------------------- | ---------------- |
| `PATIENT_ID`              | Patient identifier                                                                                                                 | Character String |
| `ENCOUNTER_ID`            | Visit identifier                                                                                                                   | Character string |
| `HOSPITAL_ENCOUNTER_DATE` | Beginning of hospital encounter date                                                                                               | Date/time        |
| `AGE_AT_VISIT`            | Patient age at admission                                                                                                           | Num              |
| `DEATH_AT_VISIT_IND`      | Indicator if the patient died during the drug overdose encounter. Leave `N/A` if patient has not died,                             | 0 /1             |
| `COUNT_CURRENT_MEDS`      | Count of active medications at the start of the drug overdose encounter                          | Num              |
| `CURRENT_OPIOID_IND`      | if the patient had at least one active medication at the start of the overdose encounter that is on the Opioids List (provided below)     | 0/1              |
| `READMISSION_90_DAY_IND`  | Indicator if the visit resulted in a subsequent readmission within 90 days     | 0/1              |
| `READMISSION_30_DAY_IND`  | Indicator if the visit resulted in a subsequent readmission within 30 days     | 0/1              |
| `FIRST_READMISSION_DATE`  | Date of the first readmission for drug overdose within 90 days. Leave `N/A` if no readmissions for drug overdose within 90 days. | Date/time        |

---

This answers our question about `DEATH_AT_VISIT_IND` earlier. Our assumption was correct to leave it as `N/A` if the patient has not died.

Now we just need to grab these fields and write it to a `CSV`.

```{r}
output <- first_readmission_dated %>%
  select(PATIENT, Id, START, AGEATSTART, DEATH_AT_VISIT_IND, COUNT_CURRENT_MEDS, CURRENT_OPOID_IND, READMISSION_90_DAY_IND, READMISSION_30_DAY_IND, FIRST_READMISSION_DATE) %>%
  rename(PATIENT_ID = PATIENT, ENCOUNTER_ID = Id, HOSPITAL_ENCOUNTER_DATE = START, AGE_AT_VISIT = AGEATSTART)
  
output
```

```{r}
readr::write_csv(output, "submission.csv", na = "N/A", append = FALSE)
```